---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# zstd

<!-- badges: start -->
[![R-CMD-check](https://github.com/r-lib/zstd/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/r-lib/zstd/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/r-lib/zstd/graph/badge.svg)](https://app.codecov.io/gh/r-lib/zstd)
<!-- badges: end -->

Zstandard compression and decompression rfor R.

## Installation

Install the stable version from CRAN:

```r
install.packages("zstd")
```

Install the development version from GitHub:

```r
pak::pak("r-lib/zstd")
```

## Usage

```{r}
library(zstd)
```

### Compress data in memory

`zstd_mem_compress()` and `zstd_mem_decompress()` work on raw vectors:

```{r}
x <- charToRaw(paste(rep("hello world ", 1000), collapse = ""))
cmp <- zstd_mem_compress(x)
length(x)
length(cmp)
identical(zstd_mem_decompress(cmp), x)
```

Use the `level` argument to trade speed for a better compression ratio:

```{r}
zstd_min_clevel()
zstd_default_clevel()
zstd_max_clevel()
cmp19 <- zstd_mem_compress(x, level = 19)
```

### Compress files

`zstd_compress()` and `zstd_decompress()` work on files:

```{r}
src <- tempfile()
writeLines(rep("hello world", 1000), src)
dst <- tempfile(fileext = ".zst")
zstd_compress(src, dst)
file.size(src)
file.size(dst)
```

`zstd_info()` shows information about compressed files:

```{r}
zstd_info(dst)[, -1]
```

```{r}
out <- tempfile()
zstd_decompress(dst, out)
identical(readLines(src), readLines(out))
```

### Compressed tar archives

`zstd_tar_compress()` creates a `.tar.zst` archive from files and
directories, and `zstd_tar_decompress()` extracts it:

```{r}
dir <- tempfile()
dir.create(file.path(dir, "subdir"), recursive = TRUE)
writeLines("hello", file.path(dir, "a.txt"))
writeLines("world", file.path(dir, "subdir", "b.txt"))

archive <- tempfile(fileext = ".tar.zst")
zstd_tar_compress(dir, archive)

exdir <- tempfile()
zstd_tar_decompress(archive, exdir)
dir(exdir, recursive = TRUE)
```

### Dictionaries

For many small, similar inputs, train a dictionary with
`zstd_train_dict()` and use it for both compression and decompression:

```{r}
samples <- lapply(1:100, function(i) {
  charToRaw(paste0('{"id":', i, ',"name":"sample"}'))
})
dict <- zstd_train_dict(samples, size = 1000)
length(samples[[1]])
length(zstd_mem_compress(samples[[1]]))
cmp <- zstd_mem_compress(samples[[1]], dict = dict)
length(cmp)
identical(zstd_mem_decompress(cmp, dict = dict), samples[[1]])
```

```{r, include = FALSE}
unlink(c(src, dst, out, dir, archive, exdir), recursive = TRUE)
```

# License

MIT © Posit Software, PBC
